---
title: "Week 1 Assignment"
author: "Xhulia Turkaj"
date: "2023-09-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment – Loading Data into a Data Frame



I have chosen to work with the dataset "weather-check.csv" from the author of  "Where People Go To Check The Weather", Walt Hickey found on the fivethirtyeight blog website. 

-Loading the packages 
```{r}
library(dplyr)
library(ggplot2)
```

-Loading the dataset from my github repository 

```{r}
weather <- read.csv("https://raw.githubusercontent.com/xhuliaturkaj/DATA-607/main/weather-check.csv",  header = TRUE, row.names = 1)

head(weather)
```



When I first run the summary function on the dataframe I realized that all the variables are of type character. I need to change that. While I change the type I will change the variable names as well. 

Starting with the first one "Do you typically check a daily weather report?" 

```{r}
weather$check_weather_daily <- as.factor(weather$Do.you.typically.check.a.daily.weather.report.)

```

And four other variables that need the same formating.
```{r}
weather$check_weather_method <-as.factor(weather$How.do.you.typically.check.the.weather.)


weather$check_weather_from_smartwatch <-as.factor(weather$If.you.had.a.smartwatch..like.the.soon.to.be.released.Apple.Watch...how.likely.or.unlikely.would.you.be.to.check.the.weather.on.that.device.)

weather$gender <-as.factor(weather$What.is.your.gender.)

weather$US.Region <-as.factor(weather$US.Region)

summary(weather$US.Region)

```

The data for the variables age and income need to be converted to ordered categorical. 

```{r}

unique(weather$Age)
# Replace "-" with NA

weather$Age[weather$Age == "-"] <- NA

#place the age groups in ascending ordered levels 
ordered_levels <- c("18 - 29", "30 - 44", "45 - 59", "60+", "NA")

weather$Age <- factor(weather$Age, levels = ordered_levels, ordered = TRUE) 

summary(weather$Age)

```

Now I will follow the same approach for the income variable. 

```{r}
unique(weather$How.much.total.combined.money.did.all.members.of.your.HOUSEHOLD.earn.last.year.)

weather$How.much.total.combined.money.did.all.members.of.your.HOUSEHOLD.earn.last.year.[weather$How.much.total.combined.money.did.all.members.of.your.HOUSEHOLD.earn.last.year.== "-"] <- NA

ordered_income_levels <- c("$0 to $9,999", "$10,000 to $24,999", "$25,000 to $49,999", "$50,000 to $74,999", "$75,000 to $99,999", "$100,000 to $124,999", "$125,000 to $149,999","$150,000 to $174,999","$175,000 to $199,999","Prefer not to answer", "NA" )

weather$income <- factor(weather$How.much.total.combined.money.did.all.members.of.your.HOUSEHOLD.earn.last.year., levels = ordered_income_levels, ordered = TRUE) 

summary(weather$income)

colnames(weather)

```

Next, I will subset the dataset to include only my variables of interest.

```{r}



weather_subset  <- na.omit(select(weather, "Age","check_weather_method", "check_weather_from_smartwatch", "check_weather_daily", "check_weather_method" ))

head(weather_subset)
```

I am interested to look into the relationship between  people's prefered method to check the weather and their age.  

```{r}
#Cross tabulation between my two variables of interest 
table(weather_subset$Age, weather_subset$check_weather_method)

```

Visualize the relationship 
```{r}

ggplot(weather_subset, aes(x = Age, fill = check_weather_method)) +
  geom_bar(position = "fill") +
  ggtitle("Preferred Method to Check Weather by Age Group") +
  xlab("Age Group") +
  ylab("Count") 


```
 
## Conclusion 
# Relationship between preferred method to check weather and age

Visually speaking, the first thing I notice is that there seems to be a negative correlation between age group and the use of the default weather app on someone's phone to check the weather. Meaning younger age groups tend to use the default weather app more than older age groups. The opposite happens with using Local TV News, where older age groups have significantly higher rates of using it to check the weather than younger age groups. 

