---
title: "Assignment 2 data 607"
author: "Xhulia Turkaj"
date: "2023-09-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment – SQL and R

 Choose six recent popular movies. Ask at least five people that you know (friends, family, classmates, imaginary
friends if necessary) to rate each of these movies that they have seen on a scale of 1 to 5. Take the results
(observations) and store them in a SQL database of your choosing. Load the information from the SQL database
into an R dataframe.

# Loading the packages

```{r}
library(DBI)
library(RMySQL)
library(tidyverse)
library(knitr)
library(tidytext)
library(wordcloud)
library(RColorBrewer)
```

# Connecting to the MySQL Database:


```{r mysql connection}


# MYSQL_PW <- readLines("pasw.txt")[1] I was attempting to resolve the pasword issue but ran out of time 




movies <- dbConnect(RMySQL::MySQL(), 
                 host = "data607xh.mysql.database.azure.com", 
                 user = "xhulia",
                 password = "Xt123456!",
                 dbname = "movie_ratings")

```

# Here I will add some SQL code chunks to recreate in this document how I set up my database

```{sql connection=movies}
DROP DATABASE IF EXISTS movie_ratings;
```


```{sql connection=movies}
CREATE DATABASE movie_ratings;
```

```{sql connection=movies}
USE movie_ratings;

```



```{sql connection=movies}
-- Create Table Movies
CREATE TABLE Movies (
    movie_id INT PRIMARY KEY AUTO_INCREMENT,
    movie_name VARCHAR(255) NOT NULL,
    release_year INT,
    overall_rating FLOAT
);
```


```{sql connection=movies}
-- Create Table Respondents
CREATE TABLE Respondents (
    respondent_id INT PRIMARY KEY AUTO_INCREMENT,
    FirstName VARCHAR(255) NOT NULL,
    LastName VARCHAR(255) NOT NULL,
    email VARCHAR(255) NOT NULL,
    UNIQUE (FirstName, LastName, email)
);
```


```{sql connection=movies}
-- Create Table ReviewsAndRatings
CREATE TABLE ReviewsAndRatings (
    rating_id INT PRIMARY KEY AUTO_INCREMENT,
    respondent_id INT,
    movie_id INT,
    review_text TEXT,
    rating INT,
    FOREIGN KEY (respondent_id) REFERENCES Respondents(respondent_id),
    FOREIGN KEY (movie_id) REFERENCES Movies(movie_id),
    CHECK (review_text IS NOT NULL OR rating IS NOT NULL)
);

```

```{sql connection=movies}
-- Insert Data into Movies
INSERT INTO Movies (movie_name, release_year) 
VALUES ('Titanic', 1997), 
       ('Barbie', 2023), 
       ('The Boy in the Striped Pajamas', 2008),
       ('Home Alone', 1990), 
       ('Mr. & Mrs. Smith', 2005),
       ('The Pod Generation', 2023);

```

```{sql connection=movies}
-- Insert Data into Respondents
INSERT INTO Respondents (FirstName, LastName, email)
VALUES ('Xhulia', 'Turkaj', 'xhulia.turkaj@example.com'),
       ('Jane', 'Smith', 'jane.smith@example.com'),
       ('Emily', 'Turkaj', 'emily.turkaj@example.com'),
       ('Michael', 'B', 'michael.b@example.com'),
       ('Sarah', 'B', 'sarah.b@example.com'),
       ('Chris', 'Chris', 'chris.chris@example.com'),
       ('Jessica', 'W', 'jessica.w@example.com'),
       ('Daniel', 'Daniel', 'daniel.daniel@example.com'),
       ('Laura', 'Walker', 'laura.walker@example.com'),
       ('Laura', 'Robinson', 'laura.robinson@example.tv');

```

```{sql connection=movies}
INSERT INTO ReviewsAndRatings (respondent_id, movie_id, review_text, rating)
VALUES
  ((SELECT respondent_id FROM Respondents WHERE email = 'xhulia.turkaj@example.com'), (SELECT movie_id FROM Movies WHERE movie_name = 'Titanic '), 'Great movie!', 5),
  ((SELECT respondent_id FROM Respondents WHERE email = 'xhulia.turkaj@example.com'), (SELECT movie_id FROM Movies WHERE movie_name = 'Barbie'), 'Not for me', 1),
  ((SELECT respondent_id FROM Respondents WHERE email = 'jane.smith@example.com'), (SELECT movie_id FROM Movies WHERE movie_name = 'Barbie'), 'Not my taste', 2),
  ((SELECT respondent_id FROM Respondents WHERE email = 'emily.turkaj@example.com'), (SELECT movie_id FROM Movies WHERE movie_name = 'The Boy in the Striped Pajamas'), 'Heartbreaking', 5),
  ((SELECT respondent_id FROM Respondents WHERE email = 'michael.b@example.com'), (SELECT movie_id FROM Movies WHERE movie_name = 'Mr. & Mrs. Smith'), 'Action-packed', 4),
  ((SELECT respondent_id FROM Respondents WHERE email = 'chris.chris@example.com'), (SELECT movie_id FROM Movies WHERE movie_name = 'Titanic '), 'Love it', NULL), 
  ((SELECT respondent_id FROM Respondents WHERE email = 'jessica.w@example.com'), (SELECT movie_id FROM Movies WHERE movie_name = 'Barbie'), NULL, 3); 
```

```{sql connection=movies}
-- Turn off Safe Updates
SET SQL_SAFE_UPDATES = 0;

```

```{sql connection=movies}
UPDATE Movies
SET overall_rating = (
    SELECT AVG(rating)
    FROM ReviewsAndRatings
    WHERE ReviewsAndRatings.movie_id = Movies.movie_id
);
```

```{sql connection=movies}
SELECT * FROM Movies;
```

```{sql connection=movies}
SELECT * FROM Respondents;
```

```{sql connection=movies}
SELECT * FROM ReviewsAndRatings;
```



# Pulling database tables  into R Data Frames

```{r}

movies_df <- dbGetQuery(movies, "SELECT * FROM Movies")
respondents_df <- dbGetQuery(movies, "SELECT * FROM Respondents")
reviews_and_ratings_df <- dbGetQuery(movies, "SELECT * FROM ReviewsAndRatings")

```

# Quick summary of the newly created dataframes
 
```{r}
summary(movies_df)
summary(respondents_df)
summary(reviews_and_ratings_df)
```
# Lets calculate again the overall rating (average) of each movie that was rated  and compare it to the overall_rating column  from the movies df which I calculated in SQL 
 
 
```{r}
# First let's join the data frames on the movie_id variable
joined_df_ratings <- inner_join(reviews_and_ratings_df, movies_df, by = "movie_id")

# Lets calculate the overall rating 

overall_ratings_df <- joined_df_ratings %>%
  filter(!is.na(rating)) %>%  
  group_by(movie_id, movie_name) %>%  
  summarise(overall_rating_R = mean(rating), .groups = "keep", na.rm = TRUE)  


print(overall_ratings_df)

print(movies_df)

summary(overall_ratings_df)
summary(movies_df)
```
# Visualize the overall rating (mean) of the  movies 
```{r}

ggplot(overall_ratings_df, aes(x = movie_name, y = overall_rating_R)) +
  geom_bar(stat = "identity") +
  labs(title = "Overall Ratings of Movies",
       x = "Movie Name",
       y = "Average Rating") 

```
# Word Cloud of the Movie Reviews

```{r}
unnested_reviews <- reviews_and_ratings_df %>%
  unnest_tokens(word, review_text)

word_count <- unnested_reviews %>%
  count(word, sort = TRUE)

wordcloud(words = word_count$word, freq = word_count$n, colors = brewer.pal(8, "Dark2"), random.order = FALSE, scale=c(3, 0.5))


```




 